The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2010-11-24 smurf_paper.tex
update notes
more detailed outline and added information about input data files

@@

\subsection{How \scuba\ takes measurements}

@@

In contrast, SCUBA-2 belongs to a new generation of cameras that uses
superconducting transition-edge sensors (TESs, like ACT, SPT,
MUSTANG). These devices are held at the critical temperature, $T_C$,
below which the material becomes a superconductor, and above which it
becomes a normal conductor. Within this transition, tiny temperature
fluctuations can produce huge changes in impedence. The devices are
operated with a voltage bias, so that thermal fluctuations produce
changing currents, which in turn generate changing local magnetic
fields. Chains of superconducting quantum interference devices
(SQUIDs) are used to amplify these signals, ultimately providing a
larger changing current that may be digitized. Like the NTD
thermistor-based cameras, SCUBA-2 is cooled in order to be background
limited. However, a superior over-all sensitivity on a per-bolometer
basis is achieved through its improved responsivity (similar to an
improved quantum efficiency for a CCD).

The load on the detectors can vary substatially, both from thermal
sources within and outside the SCUBA-2 cryostat. In order to operate
the camera across a broad dynamic range of conditions, it is therefore
necessary to thermally control the focal plane. This is achieved by
providing a fridge {\em base temperature}, $T_B$ that is well-below
$T_C$, and then using heaters that are connected to each bolometer to
place the array in to transition.

@@

\subsection{\scuba\ shared-risk observations}

@@

\begin{figure}
\centering
\includegraphics[width=0.99\linewidth]{bolos_point_mix.pdf}
\caption{A comparison between single bolometer time series in each
  band (top two plots), with the mixing chamber temperature and
  azimuth/elevation pointing offsets. There is a strong correlation
  between the bolometers and the roughly $\sim$30\,s oscillation in
  the fridge, but not with the telescope motion.}
\label{fig:bolos_mix}
\end{figure}

%------------------------------------------------------------------------------
\section{The SMURF Algorithm}
\label{sec:data}
%------------------------------------------------------------------------------

We express the signal observed by the $i$th bolometer as a function of time,
\begin{equation}
B_i(t) = G_i[E_i(t) A_i(t) + N_i(t)]
\end{equation}
where $A(t)$ is the time-varying signal produced by scanning the
telescope across astronomical sources, $E(t)$ is the time-varying
extinction, which is a function of the telescope elevation and
atmospheric conditions, and $N_i(t)$ represents sources of noise. The
two terms in square brackets, as written, have units of power
delivered to the detectors (pW). The scale factor $G_i$ converts this
effective power to the digitized units recorded on disk (DAC).

In practice, $G_i$ is further factored into two components: $G_i = R_i
D$. The responsivity, $R_i$, gives the change in current as a function
of power fluctuations (A\,pW$^{-1}$). There is then a final (and
constant across all bolometers) conversion factor from current to DAC
units, $D$ (DAC/A).

\subsection{Data cleaning}

\begin{itemize}

\item Heater ramps

\item common-mode signal amplitude

\item astronomical sources

\item flat-field source

\end{itemize}

\subsection{Correlated low-frequency noise}
At the end of the day, we iteratively fit and remove the common-mode
signal, and then any residual noise is removed with an iterative
high-pass filter. Here are the components that we know about:

\subsubsection{Fridge oscillations}
Talk about the fridge oscillation and how there might be residuals
correlated with things like the pattern of G or Tc.

@@

@@

Measure a white noise level from the power spectrum for each detector
to use as a weight. In the map, use weighted standard error to
estimate noise. This is accurate for a single pixel, but there is
still spatially correlated noise in the maps which we can show with
some figures.

@@

@@

\subsection{Bright extended emission}
The hardest thing to do with SCUBA-2 at the moment. Try to generate a
source/background mask to assist the map solution based on thresholds
in \snr.

@@

>>>>

\textbf{Guideline for figures:}

\begin{itemize}
\item minimize excess whitespace around the figure to optimize usage of space
\item ensure a decent line thickness
\item use a Times font for labels with a thickness that resembles the
  text in the body. The font size should be the same as the main body
  text or larger
\item avoid using colour unless it is necessary for emphasis
\end{itemize}

\textbf{start intro here...}

@@

My thought is that this section would give the basics about how
SCUBA-2 works, and then more specific details for the S2SRO
observations.

%--------------------------------------------------
\subsection{How \scuba\ takes measurements}
%--------------------------------------------------

@@

In contrast, SCUBA-2 belongs to a new generation of cameras that uses
superconducting transition-edge sensors (TESs, like ACT, SPT,
MUSTANG). These devices are held at the critical temperature,
$T_\mathrm{C}$, below which the material becomes a superconductor, and
above which it becomes a normal conductor. Within this transition,
tiny temperature fluctuations can produce huge changes in
impedence. The devices are operated with a voltage bias, so that
thermal fluctuations produce changing currents, which in turn generate
changing local magnetic fields. Chains of superconducting quantum
interference devices (SQUIDs) are used to amplify these signals,
ultimately providing a larger changing current that may be
digitized. Like the NTD thermistor-based cameras, SCUBA-2 is cooled in
order to be background limited. However, a superior over-all
sensitivity on a per-bolometer basis is achieved through its improved
responsivity (similar to an improved quantum efficiency for a CCD).

The load on the detectors can vary substatially, both from thermal
sources within and outside the SCUBA-2 cryostat. In order to operate
the camera across a broad dynamic range of conditions, it is therefore
necessary to thermally control the focal plane. This is achieved by
providing a fridge {\em base temperature}, $T_\mathrm{B}$ that is
well-below $T_\mathrm{C}$, and then using heaters that are connected
to each bolometer to place the array in to transition.

@@

%-------------------------------------------------
\subsection{\scuba\ shared-risk observations}
%-------------------------------------------------

@@

\begin{figure}
\centering
\includegraphics[width=0.99\linewidth]{bolos_point_mix.pdf}
\caption{A comparison between single bolometer time series in each
  band (top two plots), with the mixing chamber temperature and
  azimuth/elevation pointing offsets. There is a strong correlation
  between the bolometers and the roughly $\sim$30\,s oscillation in
  the fridge, but not with the telescope motion. Also note that the
  total power in this primary signal is similar at both 450 and
  850\,\micron, pointing to an internal rather than an external
  source.}
\label{fig:bolos_mix}
\end{figure}

\subsubsection{Fridge oscillations}
As shown in Fig.~\ref{fig:bolos_mix}, the dominant signal in both
bands is clearly an oscillation in the fridge base temperature as
evidenced by the low-resolution mixing chamber thermometer. The total
power delivered to the bolometers is comparable at each wavelength,
further suggesting a source internal to the cryostat, rather than an
external optical load. Finally, there is no obvious correspondance
with either the azimuth or elevation motion.

Show what's left over when we remove the common-mode, and the
correlation of the resulting signal with patterns in $G$ and
$T_\mathrm{C}$ across the array.

@@

\subsubsection{Line features}

The power spectra also show some narrower line features (many
correlated across the array). Even though we largely ignore them worth
some mention.

%------------------------------------------------------------------------------
\section{The SMURF Algorithm}
\label{sec:algorithm}
%------------------------------------------------------------------------------

We express the signal observed by the $i$th bolometer as a function of time,
\begin{equation}
\mathbf{b}_i(t) = g_i[\mathbf{e}_i(t) \mathbf{a}_i(t) + \mathbf{n}_i(t)]
\end{equation}
where $\mathbf{a}(t)$ is the time-varying signal produced by scanning
the telescope across astronomical sources, $\mathbf{e}(t)$ is the
time-varying extinction, which is a function of the telescope
elevation and atmospheric conditions, and $\mathbf{n}_i(t)$ represents
sources of noise. The two terms in square brackets, as written, have
units of power delivered to the detectors (pW). The scale factor $g_i$
converts this effective power to the digitized units recorded on disk
(DAC), and in this formulation is assumed to be constant in time.

In practice, $g_i$ is further factored into two components: $g_i = r_i
d$. The responsivity, $r_i$, gives the change in current as a function
of power fluctuations (A\,pW$^{-1}$). There is then a final (and
constant across all bolometers) conversion factor from current to DAC
units, $d$ (DAC/A).

As for the noise, $\mathbf{n}_i(t)$, we break it down into several
components,
%
\begin{equation}
  \mathbf{n}_i(t) = \mathbf{n}^\mathrm{w}_i(t) + \mathbf{n}^\mathrm{f}_i(t) +
  \mathbf{n}^\mathrm{c}(t),
\end{equation}
%
where $\mathbf{n}^\mathrm{w}_i(t)$ is uncorrelated white (thermal)
noise, $\mathbf{n}^\mathrm{f}_i(t)$ is low-frequency noise that is
\emph{not} correlated from bolometer-to-bolometer, and
$\mathbf{n}^\mathrm{c}(t)$ is a (predominantly low-frequency)
correlated or \emph{common-mode} component.

The primary goal of map-making is to model and remove
$\mathbf{n}^\mathrm{f}_i$ and $\mathbf{c}^\mathrm{f}_i$ from the
time-series. By re-gridding the remaining data we can then hope to
approach the theoretical noise limit in the map.

%-------------------------------------------------
\subsection{Model components}
%-------------------------------------------------

Give details about each model component, roughly in the order they are
calculated.

\subsection{Common-mode estimation}

DSB

We estimate the common-mode, remove it from the time series, use it to
flag bad bits of data, check for time-varying bolometer
responsivities... what doesn't it do?!

\subsection{Iterative filtering}

EC

The FFT filter can do high/low-pass filtering. We can also do notches
and whitening.

TJ

We also have the SMO median filter that's useful for very short scans
because edge effects don't matter so much.

@@

Measure a white noise level from the power spectrum for each detector
to use as a weight.

\subsection{Map estimation}

In the map, use weighted standard error to estimate noise. This is
accurate for a single pixel, but there is still spatially correlated
noise in the maps which we can show with some figures.

@@

%-------------------------------------------------
\subsection{Data cleaning}
%-------------------------------------------------

Since we can do most of this stuff either as a pre-processing step or
iteratively, I thinks this goes in a generic-sounding subsection. This
can go here, or perhaps earlier?

\subsubsection{Time-domain de-spiking}

DSB

DSB

\subsubsection{Gap filling / apodization}

DSB

Discuss gap filling, ``new'' window function method, also apodization
and start/end polynomial continuity stuff. Should reference CMB
literature (namely Stompor et al.).

TJ

Discuss heater ramps, normalizing off the common-mode signal
amplitude, astronomical sources, and the internal flatfield source.

\subsubsection{Additional bad bolometer rejection}

TJ

Measuring the noise power spectra and throwing away outliers. Data
flagged bad by the DA system.

\subsubsection{Map-based despiking}

EC

@@

\subsection{Bright extended emission}
The hardest thing to do with SCUBA-2 at the moment. Try to generate a
source/background mask to assist the map solution based on thresholds
in \snr. Compare with BOLOCAM stuff.

@@

Talk about potential of Pascale et al. method to get closer to
max-likelihood map solution (handle maps at constant scan angles with
weights in Fourier space separately).
